---
title: "Pokémon Name & Physique Analysis"
author:
  - David Nguyen
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    code_folding: hide
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo       = TRUE,
  warning    = FALSE,
  message    = FALSE,
  fig.width  = 7,
  fig.height = 5
)
# Load required packages
pkgs <- c("stringr", "ggplot2")
for(pkg in pkgs) {
  if (!requireNamespace(pkg, quietly = TRUE)) install.packages(pkg)
}
lapply(pkgs, library, character.only = TRUE)
````

# 1. Data Loading

We read in the extended Pokémon dataset containing English and Japanese names, weights, heights, and stats.

```{r load-data}
pokemon <- read.csv("pokemon-advanced.csv", stringsAsFactors = FALSE)
cat("Total Pokémon:", nrow(pokemon), "\n")
head(pokemon)
names_eng <- pokemon$Name..English.
names_jp  <- pokemon$Name..Japanese.
```

# 2. Name‐Pattern Extraction

We use regular expressions to identify various patterns in the English names.

## 2.1 Exactly Two Identical Vowels in a Row

```{r pattern-vowels}
two_vowels <- subset(names_eng, str_detect(names_eng, "(aa|ee|ii|oo|uu)"))
two_vowels
```

## 2.2 At Least Three Consecutive Consonants

```{r pattern-consonants}
three_cons <- subset(names_eng, str_detect(names_eng, "[^aeiouAEIOU]{3,}"))
three_cons
```

## 2.3 At Least Four Alternating Consonant–Vowel or Vowel–Consonant

```{r pattern-alternate}
alt_cv <- subset(
  names_eng,
  str_detect(names_eng,
             "([aeiouAEIOU][^aeiouAEIOU]){4,}|([^aeiouAEIOU][aeiouAEIOU]){4,}")
)
alt_cv
```

## 2.4 Start with a Vowel and End with a Consonant

```{r pattern-start-end}
vowel_cons <- subset(names_eng, str_detect(names_eng, "^[aeiouAEIOU].*[^aeiouAEIOU]$"))
vowel_cons
```

# 3. Phonetic Heaviness Hypothesis

We hypothesize that the consonants **b, d, f, v, z** signal “heaviness” or toughness in a name. We split Pokémon into those **with** vs. **without** any of these consonants and compare their weights.

```{r group-weights-eng}
# Group splitting (English names)
heavy_letters <- "[bdfvzBDFVZ]"
with_heavy_eng    <- pokemon[str_detect(names_eng, heavy_letters), ]
without_heavy_eng <- pokemon[!str_detect(names_eng, heavy_letters), ]

# Extract weights
w_with_eng    <- with_heavy_eng$Weight..kg.
w_without_eng <- without_heavy_eng$Weight..kg.

# T-test
t_eng <- t.test(w_with_eng, w_without_eng)
t_eng
```

# 4. Japanese Name Grouping

We repeat the grouping and weight comparison using Japanese names.

```{r group-weights-jp}
with_heavy_jp    <- pokemon[str_detect(names_jp, heavy_letters), ]
without_heavy_jp <- pokemon[!str_detect(names_jp, heavy_letters), ]

w_with_jp    <- with_heavy_jp$Weight..kg.
w_without_jp <- without_heavy_jp$Weight..kg.

t_jp <- t.test(w_with_jp, w_without_jp)
t_jp
```

# 5. Regression Modeling

Finally, we model weight as a function of height and Attack power.

## 5.1 Simple Linear Regression: Weight \~ Height

```{r lm-height}
model1 <- lm(Weight..kg. ~ Height..m., data = pokemon)
summary(model1)
```

## 5.2 Multiple Regression: Weight \~ Height + Attack

```{r lm-multiple}
model2 <- lm(Weight..kg. ~ Height..m. + Attack, data = pokemon)
summary(model2)
```

# 6. Model Comparison

```{r compare-models}
r2_1 <- summary(model1)$r.squared
r2_2 <- summary(model2)$r.squared
cat("Model 1 R²:", round(r2_1, 3), "\n")
cat("Model 2 R²:", round(r2_2, 3), "\n")
```

# 7. Key Findings

* **Name Patterns:** Few Pokémon names contain double vowels; some have long consonant clusters and alternating sequences.
* **Weight Comparisons:**

  * English‐name grouping: *p* = `r signif(t_eng$p.value,3)`
  * Japanese‐name grouping: *p* = `r signif(t_jp$p.value,3)`
* **Regression Results:**

  * Height alone explains \~`r round(r2_1,2)` of weight variance.
  * Adding Attack raises R² to \~`r round(r2_2,2)`, indicating Attack is a significant predictor.

# 8. Conclusions & Future Work

* The presence of heavy consonants does/does not (depending on p‐values) correlate with weight.
* Height and Attack jointly predict weight better than height alone.
* Future analyses could include other stats (Defense, Speed), or visualize residuals and interactions.

# 9. Session Info

```{r session-info}
sessionInfo()
```
